---
title: "ST558 Project 2: Creating predictive models and automating Markdown reports."
author: "Josh Baber & Lan Lin"
date: '2022-07-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'F:\\Graduate\\NCSU_courses\\ST558\\projects\\Project_2\\OnlineNewsPopularity')
```

```{r, include=FALSE}
library(tidyverse)
library(reader)
library(corrplot)
library(ggridges)
library(RColorBrewer)
library(GGally)
```


Read in the data set
```{r}
# Read in the data set and remove the non-predictive variables 
shares_Data <- read_csv("OnlineNewsPopularity.csv")[-2:-1]
head(shares_Data)
```

Convert the dummy variables of channels to single categorical variable
```{r}
# create a single variable representing the data channel
channel <- factor(cbind(VALUE = factor(max.col(shares_Data[12:17]), ordered = TRUE)))
levels(channel) <- c( 'Lifestyle', 'Entertainment', 'Business', 'Social Media', 'Tech', 'World')

# Create a new data set using the single variable representing the data channel
shares_Data_chl <- shares_Data %>% select(-starts_with("data_channel")) %>% 
                     mutate(channel) %>% 
                     select(channel, everything())
```

Convert the dummy variables of weekdays to single categorical variable
```{r}
# create a single variable representing the data weekdays
weekday <- factor(cbind(VALUE = factor(max.col(shares_Data_chl[25:31]), ordered = TRUE)))
levels(weekday) <- c("Monday", 'Tuesday',' Wednesday', 'Thursday',' Friday', 'Saturday', 'Sunday')

# Create a new data set using the single variable representing the data channel
shares_Data_chl <- shares_Data_chl %>% select(-starts_with("weekday_is")) %>% 
                     mutate(weekday) %>% 
                     select(channel, weekday, everything())

shares_Data_chl
```

```{r}
names(shares_Data_chl)
```

Tuesday Wednesday Thursday Friday Saturday Sunday


Subset the data to work on the "Lifestyle" data channel 
```{r}
shares_Lifestyle <- shares_Data_chl %>% filter(channel == "Lifestyle")
```

Statistical summary of response variable shares 
```{r}
summary(shares_Lifestyle$shares)
```

Boxplot of shares
```{r}
ggplot(shares_Lifestyle, aes(y= shares)) + geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim = quantile(shares_Lifestyle$shares, c(0.1, 0.9)))
```

Distribution of Shares by weekday using the shares_Lifestyle data set. The grouping variable weekday will be mapped to the y-axis:
```{r}
ggplot(shares_Lifestyle, 
       aes(x = shares, 
           y = weekday)) +
  geom_density_ridges(aes(fill = weekday)) + 
  coord_cartesian(xlim = quantile(shares_Lifestyle$shares, c(0, 0.9))) +
  theme_ridges() +
  labs("Highway mileage by auto class") +
  theme(legend.position = "none")
```

```{r}
names(shares_Lifestyle)[c(2:5, 8, 10:12, 16, 19, 22, 25, 32:35, 38, 41, 44:48)]
```

Data set for modeling 
```{r}
Lifestyle <- shares_Lifestyle[c(2:5, 8, 10:12, 16, 19, 22, 25, 32:35, 38, 41, 44:48)]
Lifestyle
```


```{r}
Correlation <- cor(shares_Lifestyle[c(3:5, 8, 10:12, 16, 19, 22, 25, 32:35, 38, 41, 44:48)], method = "spearman")
Correlation
```

```{r, fig.height= 12, fig.width = 18}
corrplot(Correlation, type = "upper", tl.pos = "lt")
corrplot(Correlation, type = "lower", method = "number", add = TRUE, diag = FALSE, tl.pos = "n")
```